---
title: "p8105_hw5_ha2546"
author: "Hana Akbarnejad"
date: "11/3/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

```{r import,replace_NA}

head(iris, 5)

set.seed(10)

iris_with_missing = iris %>% 
  map_df(~replace(.x, sample(1:150, 20), NA)) %>%
  mutate(Species = as.character(Species))
```

_iris_ is a dataset that contains `r nrow(iris)` observations of `r n_distinct(pull(iris, Species))` different flower species: `r unique(pull(iris, Species))`. It stores information in `r ncol(iris)` variables about flowers: species, sepal length and width, and petal length and width.
First, I chose a sample of 20 from these 150 obserations, and eplaced them with NA. I also converted the species factor variable to numeric. 

```{r fill_NA}


replace_NA_fnc = function(vect){
  
  if(is.character(vect)){
    replace_na(vect, "virginica")
  } else if (is.numeric(vect)) {
    replace_na(vect, mean(vect, na.rm = TRUE))
  }
  
}


iris_output = map(iris_with_missing, replace_NA_fnc) %>%  as_tibble() 
```

